High-Dimensional Regularized Discriminant Analysis

### 2013-12-26 reports/paper-hdrda.Rnw
Corrected derivation of cov estimator's interpretation.
End result was correct but intermediate steps were not
Rewrote the abstract and the introduction.
- The intro is a bit verbose and needs editing.
- Related to #10 and #15.

LDA and QDA suck. Friedman to the rescue with RDA. We reformulate the RDA
classifier. Our parameterizaton yields 1) improved interpretation, 2) nice
properties that facilitate dimension reduction, and 3) as a result of the
dimension reduction, efficient calculation for high-dimensional data.

Here, we propose a new regularized discriminant analysis classifier that
incorporates dimension reduction and covariance-matrix shrinkage to enable a
computationally efficient classifier. We demonstrate with four high-dimensional
microarray data sets that our proposed classifier is often superior to several
recently proposed sparse and regularized classifiers in terms of classification
accuracy. Finally, we provide an implementation of our proposed classifier in
the {\tt sparsediscrim} R package, available on CRAN.

@@

Linear discriminant analysis (LDA) and quadratic discriminant analysis (QDA) are
classical approaches to supervised learning and are widely used in a variety of
areas, such as medical research, facial recognition, and TODO
\cite{Dudoit:2002ev,TODO1,TODO2} Both the LDA and QDA classifiers are based on
the assumption that the underlying distribution of each of the $K$ classes are
multivariate normally distributed. While both methods are popular in practice,
they often cannot be applied to modern data sets because the sample covariance
matrices employed are singular when the number of features exceeds the total
sample size $N$. For instance, standard benchmark microarray data sets used in
biological research often consist of between $p = 10,000$ and $55,000$ probes
and fewer than $N = 100$ observations, in which case LDA and QDA are
incalculable.

\cite{Friedman:1989tm} proposed regularized discriminant analysis (RDA), which
partially pools the individual sample covariance matrices used in the QDA
classifier with the pooled sample covariance matrix from the LDA
classifier. After pooling the covariance matrices, the RDA classifier employs a
covariance-matrix regularization technique that shrinks the zero eigenvalues of
the resulting covariance matrices to ensure they are nonsingular. Similarly,
several researchers, including \cite*{Zhang:2010va}, \cite{Ye:2009gd},
\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
\cite{Ye:2006vx}, and \cite{Ye:2005uu}, have proposed regularization methods to
improve the estimation of the population covariance matrices before applying
either the LDA or QDA classifiers. Covariance-matrix regularization methods
often are based on the well-known ridge-regression approach of
\cite{Hoerl:1970cd} and adjust the eigenvalues of the sample covariance matrices
to ensure positive definiteness. For excellent overviews of the regularization
problem in discriminant analysis, see \cite{Murphy:2012uq},
\cite{Mkhadri:1997gy}, and \cite{Friedman:1989tm}.

% TODO: Should we remove this paragraph? Keep a couple of the recent references
% though.
Alternatively, many researchers, including \cite{Mai:2012bf}, \cite{Fan:2012iq},
\cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, and \cite{Huang:2010ju}, have
introduced restrictions on the population covariance matrices in linear and
quadratic discriminant analysis to reduce the number of covariance parameters
that must be estimated to ensure that the resulting classifiers are
calculable. For instance, \cite{Dudoit:2002ev} have modified the linear and
quadratic discriminant analysis classifiers by assuming that the population
covariance matrices are diagonal, resulting in classifiers that are special
cases of the Naive Bayes classifier \citep[Chapter 4]{Hastie:2008dt}.

We reformulate the RDA classifier. Our parameterizaton yields 1) improved
interpretation, 2) nice properties that facilitate dimension reduction, and 3)
as a result of the dimension reduction, efficient calculation for
high-dimensional data.

Here, we propose the high-dimensional regularized discriminant analysis
(\emph{HDRDA}) classifier by first constructing a pooling matrix as a convex
combination of the covariance-matrix estimators employed in linear and quadratic
discriminant analysis. Given that the assumption of equal population covariance
matrices for linear discriminant analysis is often too stringent in practice
\citep{Clemmensen:2011kr}, our proposed classifier can relax the linearity
assumption by instead incorporating unequal covariance matrices while retaining
linear and quadratic discriminant analysis as special cases. We then establish
the \emph{HDRDA} classifier by applying covariance-matrix regularization with a
parameterization that includes a large family of shrinkage methods, such as
those from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods
studied by \cite{Ramey:2013ji} and \cite{Xu:2009fl}. This regularization has the
effect of stabilizing the eigenvalues of the covariance-matrix estimators and
ensuring that the classifier is well-posed.

We show that our proposed classifier is invariant to any adjustments of the
eigenvalues corresponding to the null space of the pooled sample covariance
matrix. Applying a reasoning similar to \cite{Ye:2006tq}, we decompose the
decision rule of the \emph{HDRDA} classifier and show that the sum related to
the null space of the pooled sample covariance matrix can be discarded. Hence,
no gain in classification performance can be obtained by adjusting these zero
eigenvalues. This invariance property provides a substantial increase in
computational efficiency of model training and selection. In fact, we show that
the inverse and determinants can be reduced to calculations of $n_k \times n_k$
matrices, where $n_k$ is the sample size of the $k$th class, $k = 1, \ldots, K$.
This reduction in calculation is clearly advantageous because $n_k$ is often
quite small in microarray data sets (e.g., 10--20 observations).

We also demonstrate that the \emph{HDRDA} classifier is often superior to
several recent classifiers designed for small-sample, high-dimensional data. In
particular, we compare our proposed classifier with diagonal linear discriminant
analysis proposed by \cite{Dudoit:2002ev} and the recently proposed classifiers
from \cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
\cite{Witten:2011kc}. We conduct our comparison of the classification accuracy
of the competing classifiers using four small-sample, high-dimensional benchmark
microarray data sets.

\subsection{Preliminaries}
To facilitate our discussion of covariance-matrix regularization and regularized
discriminant analysis, we require the following notation. Let $\mathbb{R}_{a
  \times b}$ denotes the matrix space of all $a \times b$ matrices over the real
field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity matrix, and
let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm
0_m$ is understood to denote $\bm 0_{m \times m}$. Let $\bm A \oplus \bm B$
denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in
\mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^{T}$,
$\bm A^+$, and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A
\bm z = \bm 0_{m \times 1}\}$ denote the transpose, the Moore-Penrose
pseudoinverse, and the null space of $\bm A \in \mathbb{R}_{m \times p}$,
respectively. Let $\mathbb{R}_{p \times p}^{>}$ denote the cone of real $p
\times p$ positive-definite matrices. Similarly, let $\mathbb{R}_{p \times
  p}^{\ge}$ denote the cone of real $p \times p$ positive-semidefinite
matrices. Let $V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y^{T} \bm v
= 0$ for all $\bm v \in V \}$ denote the orthogonal complement of a vector space
$V \subset \mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if
$c \ne 0$ and $0$ otherwise.

@@

\subsection{Model Specification}

\cite{Halbe:2007ea} demonstrate that  parameterization results in a

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
1, \ldots, N)$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda) &= \widehat{\bm \Sigma}_k + \lambda (\widehat{\bm \Sigma} - \widehat{\bm \Sigma}_k) \nonumber\\
%	&= \widehat{\bm \Sigma}_k + \lambda \left(\sum_{k' = 1}^K \frac{n_k}{N} \widehat{\bm \Sigma}_{k'}  - \widehat{\bm \Sigma}_k \right) \nonumber\\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}} n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
%	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N I(y_i = k) x_i x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k) x_i x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$. We
can interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix
estimator that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to
better estimate $ \bm \Sigma_k$. From \eqref{eq:sig-lambda-alternative2},
$\lambda$ weights the contribution of each of the $N$ observations in estimating
$\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
$N$ rather than $n_k$ observations. Below, we will show that the pooling
operation is advantageous in increasing the rank of each $\widehat{\bm
  \Sigma}_k(\lambda)$ from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm
  \Sigma})$ for $0 < \lambda \le 1$. Notice that if $\lambda = 0$, then the
observations from the remaining $K - 1$ classes do not contribute to the
estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm
  \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in
\eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
      [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$
      such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for
      $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
        \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k =
      \bm X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the
      contours of five multivariate normal populations for $\lambda = 0$ with
      unequal covariance matrices. As $\lambda$ approaches 1, the contours
      become more similar, resulting in identical contours for $\lambda = 1$.
\begin{figure}
<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 1/4)
rda_contours(lambda = 3/4)
rda_contours(lambda = 1)
@ 
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:hdrda-contours}
\end{figure}
As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
\emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We
provide an implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the
{\tt sparsediscrim} R package, available on CRAN.

\subsection{Relationship with Regularized Discriminant Analysis}

@@

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.

\section{Properties of the HDRDA}

\subsection{Enhanced Computational Efficiency}
\label{sec:computational-efficiency}

>>>>

The popular regularized discriminant analysis (RDA) classifier proposed by
\citeauthor{Friedman:1989tm} utilizes a biased covariance-matrix estimator that
partially pools the sample covariance matrices from linear and quadratic
disriminant analysis before shrinking the resulting estimator towards a scaled
identity matrix. The RDA classifier's two tuning parameters are typically
estimated via a computationally burdensome cross-validation procedure from a
grid of candidate values. We reformulate the RDA classifier in the
high-dimensional setting and then show that the classification decision rule is
equivalent to one in a subspace having a much lower dimension. As a result, the
dimension reduction involved yields a substantial reduction in computation
during model selection. Our parameterization also offers interpretability
previously lacking with the RDA classifier. We demonstrate with three artificial
data sets and four high-dimensional real data sets that our proposed classifier
is often superior to several recently proposed sparse and regularized
classifiers in terms of classification accuracy. Finally, we provide an
implementation of our proposed classifier in the {\tt sparsediscrim} R package,
available on CRAN.

@@

As machine-learning algorithms are increasingly utilized to automate data-driven
decisions, researchers continue to develop a wealth of \emph{classifiers} to
construct automated pattern-recognition systems in a variety of disciplines,
such as medical research, facial recognition, remote sensing, and chemical
analysis \citep{TODO}. Despite the advances in classification to facilitate
technological advancements, well-established classifiers often degrade in
accuracy with modern, high-dimensional data sets when the number of features $p$
exceeds the total sample size $N$, due to the \emph{curse of dimensionality}
\citep{Bellman:1961tn}. For instance, linear discriminant analysis (LDA) and
quadratic discriminant analysis (QDA) are popular because of their simplicity,
speed of computation, and interpretability. However, both methods are
incalculable when $p > N$ because the class sample covariance matrices are
singular \citep*{Bouveyron:2007gx}.

As a result, researchers have emphasized three standard approaches to modifying
the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
data: 1) covariance-matrix regularization, 2) dimension reduction, and 3) model
parsimony. Covariance-matrix regularization methods often are based on the
well-known ridge-regression approach of
\cite{Hoerl:1970cd}. \cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd},
\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
\cite{Ye:2006vx}, and \cite{Ye:2005uu} have proposed regularization methods to
improve the estimation of the class covariance matrices by adjusting the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. Alternatively, dimension reduction typically decomposes a model to
orthogonal components and retains the components corresponding to the model's
intrinsic dimension. While some classificatory information may be lost,
dimension reduction methods have been shown to be effective in practice
\citep{YePaper,TODO}. Finally, \cite{Mai:2012bf}, \cite{Fan:2012iq},
\cite{Merchante:2012vk}, \cite{Clemmensen:2011kr}, \cite{Huang:2010ju}, and
\cite{Dudoit:2002ev} have proposed restrictions on the population covariance
matrices to reduce the number of parameters that must be estimated. For
instance, \cite{Dudoit:2002ev} have assumed the features within each class are
independent so that the population covariance matrices are diagonal, resulting
in positive-definite sample covariance matrices. For excellent overviews of the
regularization problem in discriminant analysis and other techniques, see
\cite{Murphy:2012uq} and \cite{Mkhadri:1997gy}.

We consider regularized discriminant analysis (\emph{RDA}) from a classic paper
by \cite{Friedman:1989tm}, who proposed improvements to the \emph{QDA}
classifier by employing a covariance matrix estimator that partially pools the
individual sample covariance matrices with the pooled sample covariance matrix
from the \emph{LDA} classifier. Given that the assumption of equal population
covariance matrices for the \emph{LDA} classifier is often too stringent in practice
\citep{Clemmensen:2011kr}, the \emph{RDA} classifier can relax the linearity
assumption by instead incorporating unequal covariance matrices while retaining
as special cases the \emph{LDA} and QDA classifiers. Covariance-matrix
regularization is then applied to the partially pooled covariance-matrix
estimator by shrinking it towards a scaled identity matrix to ensure positive
definiteness. We argue that the \emph{RDA} classifier inherently employs
dimension reduction in addition to regularization. After parameterizing the
\emph{RDA} classifier similar to \cite{Hastie:2008dt} and \cite{Halbe:2007ea},
we demonstrate that the intrinsic dimension of the \emph{RDA} classifier is
equal to the training sample size of each class.\footnote{TODO: Introduce HDRDA
  somewhere.} We then show that the \emph{RDA} decision rule is invariant to
adjustments of the zero eigenvalues. We show that our proposed classifier is
invariant to any adjustments of the eigenvalues corresponding to the null space
of the pooled sample covariance matrix. Applying a reasoning similar to
\cite{Ye:2006tq}, we decompose the decision rule of the \emph{HDRDA} classifier
and show that the sum related to the null space of the pooled sample covariance
matrix can be discarded. Hence, no gain in classification performance can be
obtained by adjusting these zero eigenvalues. This invariance property provides
a substantial increase in computational efficiency of model training and
selection. Moreover, the decision rule in the original feature space is
equivalent to a decision rule in a much lower dimension (i.e., the intrinsic
dimension), such that the matrix inverses and determinants can be rapidly
computed. This invariance property as well as the dual optimization problem
result in substantial reduction in computation. Additionally, our
parameterization of the \emph{RDA} classifier offers interpretability that has
been previously lacking \citep{Bensmail:1996fw}. We explicitly show that the
pooling parameter in the \emph{RDA} covariance-matrix estimator controls the
contribution of each training observation to the estimation of each class
covariance matrix. Furthermore, our parameterization includes a large family of
shrinkage methods, such as those from \cite{Srivastava:2007ww},
\cite{Rao:1971ul}, and other methods studied by \cite{Ramey:2013ji} and
\cite{Xu:2009fl}.

Given the \emph{RDA} classifier has been shown to have excellent performance in
the high-dimensional setting \citep{Webb:2011vu} but is limited by its
computationally intense model selection procedure, our work enables the
\emph{RDA} classifier to be employed with high-dimensional data in
practice. This is reassuring because the \emph{RDA} classifier remains widely
popular in the literature. In fact, variants of the RDA classifier have been
applied to microarray data \citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te},
facial recognition
\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
spectra \citep{Wu:1996us,Aeberhard:1993fp}.

After demonstrating that the \emph{HDRDA} classifier can be computed efficiently
and establishing its improved interpretability, we compare the computational
runtime of the \emph{HDRDA} classifier with an implementation of the \emph{RDA}
classifier from the {\tt klaR} R package to illustrate that the \emph{HDRDA}
classifier greatly reduces computation when $p \gg N$. We then study the
classification performance of the \emph{HDRDA} classifier on several artificial
and real high-dimensional data sets. We compare the \emph{HDRDA} classifier with
several recent several recent classifiers from \cite{Dudoit:2002ev},
\cite{Guo:2007te}, \cite{Pang:2009ik}, \cite{Tong:2012hw}, and
\cite{Witten:2011kc}, each of which is designed for small-sample,
high-dimensional data. We demonstrate that the \emph{HDRDA} classifier often
attains superior classification accuracy. We also include in our classification
study a support vector machine (\emph{SVM}) and $k$-nearest neigbors ($kNN$) as
benchmarks.

The remainder of this document is organized as follows. In Section
\ref{sec:preliminaries} we presents the discrimination problem and the necessary
notation to describe our contributions. In Section \ref{sec:rda} we present our
\emph{HDRDA} classifier along with its interpretation, properties, and
computationally efficient model-selection procedure. Section \ref{sec:sims}
describes our simulation studies of artificial and real data sets and also
describes the experimental results. We conclude with a brief discussion in
Section \ref{sec:discussion}.

\section{Preliminaries}
\label{sec:preliminaries}

\subsection{Notation}

To facilitate our discussion of covariance-matrix regularization and regularized
discriminant analysis, we require the following notation. Let $\mathbb{R}_{a
  \times b}$ denote the matrix space of all $a \times b$ matrices over the real
field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity matrix, and
let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm
0_m$ is understood to denote $\bm 0_{m \times m}$. Let $\bm A \oplus \bm B$
denote the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in
\mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^{T}$,
$\bm A^+$, and $\mathcal{N}(\bm A) = \{\bm z \in \mathbb{R}_{p \times 1} : \bm A
\bm z = \bm 0_{m \times 1}\}$ denote the transpose, the Moore-Penrose
pseudoinverse, and the null space of $\bm A \in \mathbb{R}_{m \times p}$,
respectively. Denote by $\mathbb{R}_{p \times p}^{>}$ the cone of real $p \times
p$ positive-definite matrices. Similarly, let $\mathbb{R}_{p \times p}^{\ge}$
denote the cone of real $p \times p$ positive-semidefinite matrices. Let
$V^{\perp} = \{\bm y \in \mathbb{R}_{p \times 1} : \bm y^{T} \bm v = 0$ for all
$\bm v \in V \}$ denote the orthogonal complement of a vector space $V \subset
\mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$
and $0$ otherwise.

\subsection{Discriminant Analysis}

@@

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) = (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
1, \ldots, N)$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is, from
\eqref{eq:pooled-cov} we have
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$. We
can interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix
estimator that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to
better estimate $ \bm \Sigma_k$. From \eqref{eq:sig-lambda-alternative2},
$\lambda$ weights the contribution of each of the $N$ observations in estimating
$\bm \Sigma_k$. Hence, for $0 < \lambda \le 1$, we estimate $\bm \Sigma_k$ with
$N$ rather than $n_k$ observations.\footnote{TODO: Add more to the
  interpretation here.}

\begin{figure}
<<rda_weights>>=
n <- c(50, 25, 15, 10)
N <- sum(n)
lambda <- seq(0, 1, length = 1000)

rda_weights <- sapply(seq_along(n), function(k) {
      lambda / N + (1 - lambda) / n[k]
      })
rda_weights <- data.frame(lambda = lambda, rda_weights)
rda_weights <- melt(rda_weights, id = 'lambda')
colnames(rda_weights) <- c("lambda", "Class", "Weight")
levels(rda_weights$Class) <- seq_along(n)

p <- ggplot(rda_weights, aes(lambda, Weight, linetype = Class, color = Class))
p <- p + geom_line(size = 1)
p <- p + theme_bw() + xlab(expression(lambda))
p <- p + theme(axis.text = element_text(size = 10))
p <- p + theme(strip.text.x = element_text(size = 16))
p <- p + theme(legend.title = element_text(size = 16, face = "bold"))
p <- p + theme(legend.text = element_text(size = 15))
p <- p + theme(legend.key.size = unit(1, "cm"))
p
@ 
\caption{Weights $c_{ik}(\lambda)$ of \emph{RDA} classifier as a function of the
  pooling parameter $\lambda$ when $n_1 = 50$, $n_2 = 25$, $n_3 = 15$, and $n_4$
  = 10.}
\label{fig:hdrda-weights}
\end{figure}

Below, we will show that the pooling
operation is advantageous in increasing the rank of each $\widehat{\bm
  \Sigma}_k(\lambda)$ from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm
  \Sigma})$ for $0 < \lambda \le 1$. Notice that if $\lambda = 0$, then the
observations from the remaining $K - 1$ classes do not contribute to the
estimation of $\bm \Sigma_k$, corresponding to $\widehat{\bm
  \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights in
\eqref{eq:sig-lambda-alternative2} reduce to $1/N$, corresponding to
$\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we define $\bm X =
      [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm x_N^{T}]^{T}$
      such that $\widehat{\bm \Sigma} = \bm X^{T} \bm X$. Similarly, for
      $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
        \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k =
      \bm X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the
      contours of five multivariate normal populations for $\lambda = 0$ with
      unequal covariance matrices. As $\lambda$ approaches 1, the contours
      become more similar, resulting in identical contours for $\lambda = 1$.
\begin{figure}
<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 1/4)
rda_contours(lambda = 3/4)
rda_contours(lambda = 1)
@ 
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:hdrda-contours}
\end{figure}
As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k = \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
\emph{RDA} classifier from \cite{Friedman:1989tm}. Now, substituting
\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation \citep{Hastie:2008dt}. We
provide an implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the
{\tt sparsediscrim} R package, available on CRAN.

\subsection{Relationship with Friedman's RDA Classifier}

@@

To further improve the estimation of $\bm \Sigma_k$ and to stabilize the inverse
of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has proposed the biased
covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
established by substituting \eqref{eq:sig-rda} into
\eqref{eq:qda}.\footnote{TODO: Discuss $\alpha_k \equiv 1-\gamma$, etc.}

\subsection{Properties of the HDRDA Classifier}
\label{sec:computational-efficiency}
